---
title: "DSC630Assignment2BKudaimiRMD"
author: "Bilal Kudaimi"
date: "6/20/2021"
output: pdf_document
---

## The goal of this assignment is to find out through EDA and regression which days are the best for running a marketing campaign to increase game attendee number. We have a dataset of LA Dodgers games with information such as month, day, daily temperature, game opponent, weather, number of attendees, and whether items such as caps, shirts, fireworks, and bobbleheads are sold at the game. 

## We will run the marketing campaign on days with the highest attendee number to increase the campaign's audience. To find out which days have the highest attendee number, we will use regression to tell us which features of this dataset contribute the most to attendee number, and from this, we can tell what days (e.g. days in March, Saturdays, etc) have the most weight in the model, and thus, which days would be the best to run our campaign.

## EDA will be conducted to gain insight into the data, then multiple regression will be conducted to find out which features of the dataset weigh the most in predicting game attendee number.

## Importing the data and viewing structure and summary statistics

```{r code, echo = FALSE}
#The goal of this assignment is to find out through EDA and regression which days 
#are the best for running a marketing campaign to increase game attendee number.
#We will run the marketing campaign on days with the highest attendee number to increase
#its audience. Regression will tell us which features of this dataset contribute the most 
#to attendee number, and from this, we can tell what days would be the best to run our campaign.

#EDA will be conducted to gain insight into the data, then multiple regression will 
#be conducted to find out which features of the dataset weigh the most in predicting game attendee number.

#Setting the working directory
setwd("C:/Users/PS3ma/Documents/Bellevue University/DSC 630")

#Importing the necessary libraries
library(Hmisc)
library(ggplot2)
library(QuantPsyc)
#Importing the data
dodgers <- read.csv('dodgers.csv')

#Checking if there is any missing data
is.null(dodgers)


#Viewing the data structure and summary statistics
str(dodgers)
describe(dodgers)
```

## Generating histograms and bar charts of the continuous and categorical variables, respectively. Boxplots will also be generated for the numerical variables. This will reveal if there is any skewness among the variables.

```{r part2, echo = FALSE}

#Generating histograms and bar charts of the key variables.
#In addition to histograms, boxplots of the numerical variables will be generated.
#This will reveal if there is any skewness among the variables.
ggplot(dodgers, aes(day)) + geom_histogram()
ggplot(dodgers, aes(attend)) + geom_histogram()
ggplot(dodgers, aes(temp)) + geom_histogram()
ggplot(dodgers, aes(opponent)) + geom_bar()
ggplot(dodgers, aes(skies)) + geom_bar()
ggplot(dodgers, aes(day_night)) + geom_bar()
ggplot(dodgers, aes(cap)) + geom_bar()
ggplot(dodgers, aes(shirt)) + geom_bar()
ggplot(dodgers, aes(fireworks)) + geom_bar()
ggplot(dodgers, aes(bobblehead)) + geom_bar()
ggplot(dodgers, aes(temp)) + geom_boxplot()
ggplot(dodgers, aes(attend)) + geom_boxplot()
ggplot(dodgers, aes(day)) + geom_boxplot()
```

## There does not appear to be any skewness, however, most of the games were played on clear nights and souvenirs such as caps, shirts, fireworks, and bobbleheads weren't sold at a majority of games. It also appears the 15th of each month is a popular day for attending baseball games, for some reason.

## Two scatter plots will now be generated of game attendees vs temperature and day of the month to determine if temperature or day number affects the number of attendees.

```{r part3, echo = FALSE}
#There does not appear to be any skewness, however, most of the games were played on clear nights 
#and souvenirs such as caps, shirts, fireworks, and bobbleheads weren't sold at a majority of games

#Generating a scatter plot of game attendees vs temperature and day of the month to see if 
#temperature or day affects the number of attendees
ggplot(dodgers, aes(temp, attend)) + geom_point()
ggplot(dodgers, aes(day, attend)) + geom_point()
```

## It doesn't look like temperature or day number is correlated with number of attendees, so the other variables will be plotted against attendee number to see if there are any correlations. Boxplots will be used since the remaining variables to be plotted are categorical in nature.

```{r part4, echo = FALSE}
#It doesn't look like temperature or day number is correlated with number of attendees, so the other variables will be
#plotted against attendee number to see if there are any correlations. Boxplots will be used since the remaining 
#variables to be plotted are categorical in nature.
ggplot(dodgers, aes(skies, attend)) + geom_boxplot()
ggplot(dodgers, aes(opponent, attend)) + geom_boxplot()
ggplot(dodgers, aes(day_night, attend)) + geom_boxplot()
ggplot(dodgers, aes(cap, attend)) + geom_boxplot()
ggplot(dodgers, aes(shirt, attend)) + geom_boxplot()
ggplot(dodgers, aes(fireworks, attend)) + geom_boxplot()
ggplot(dodgers, aes(bobblehead, attend)) + geom_boxplot()
ggplot(dodgers, aes(day_of_week, attend)) + geom_boxplot()
```

## It appears that Tuesdays have the highest rate of attendance, and that bobbleheads and shirts present at a game are correlated with high attendance. It also appears that games vs the Angels, Mets, and Nationals garner the highest median attendee number, while games vs the Braves and the Pirates garner the lowest median attendee number. Based on the EDA, it appears that games on Tuesdays vs the Angels, Mets, or Nationals where shirts and bobbleheads are sold would be a good target for a marketing campaign.

## Multiple linear regression will now be set up to determine which of the factors weigh into game attendee number the most. Any days of the week identified as having a significant weight in attendee number will be reported. First, though, it must be determined that none of the numeric variables are correlated with each other. Variables that have correlations with each other, or that have no correlation with the target variable (attendee number) will be dropped.

```{r part5, echo = FALSE}
#It appears that Tuesdays have the highest rate of attendance, and that bobbleheads and shirts present 
#at a game are correlated with high attendance. It also appears that games vs the Angels, Mets, and Nationals 
#garner the highest median attendee number, while games vs the Braves and the Pirates garner the lowest median 
#attendee number. Based on the EDA, it appears that games vs the Pirates and Braves where shirts and bobbleheads
#are sold would be a good target for a marketing campaign.

#Multiple linear regression will now be set up to determine which of the factors weigh into game attendee number the #most. Any days of the week identified as having a significant weight in attendee number will be reported. 
#First, though, it must be determined that none of the numeric variables are correlated with each other.
#Variables that have correlations with each other, or that have no correlation with the target variable (attendee #number) will be dropped.
cor(dodgers[sapply(dodgers,is.numeric)])
```

## None of the numeric features appear to be correlated with each other, so we can continue with building the model. Since temperature, day number, and skies did not appear to be correlated with attendee number, these variables will not be used in the regression model.

```{r part6, echo = FALSE}
#None of the numeric features appear to be correlated with each other, so we can continue with building the model. 
#Since temperature, day number, and skies did not appear to be correlated with attendee number, these variables 
#will not be used in the regression model.

model <- lm(attend ~ month + day_of_week + opponent + cap + shirt + fireworks + bobblehead, data = dodgers)
summary(model)

#Based on the results of the regression model summary, the days with the highest attendee number are Sundays and Tuesdays.
#In addition, any day when bobbleheads are being sold have high attendee number. This matches the results of the EDA, 
#confirming that these are the best days to run our marketing campaign.
```

## The model coefficients and significance codes will reveal how much weight each variable has in predicting attendee numbers. Based on these coefficients in the regression model summary, the days with the highest attendee number are Sundays and Tuesdays. In addition, any day when fireworks and bobbleheads are being sold have high attendee numbers. This matches the results of the EDA, although opponent did not weigh heavily into the regression model. In addition, the multiple R-squared value was 0.703, meaning that this model is a good fit for this data, and the p-value was 3.7e-5, meaning it is highly unlikely this model fit the data by chance. It is confirmed, then, that the best days to run our marketing campaign to spread it to a wider audience are Sundays, Tuesdays, and any day when fireworks and bobbleheads are sold.